---
title: "Data Analysis for the exam in Spatial Analytics"
author: "Aske Svane Qvist & Hania Matera"
date: "10 June 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Analysis 

This script contains data pre-processing, visualisations, as well as spatial analysis for our exam project on the spatial influence on coffee production.

- The longitude and latitude coordinates in the csv-file have been manually annotated using Google Maps.

## Packages
```{r}
library(pacman)
p_load(tidyverse, sf, mapview, spdep, leaflet, worldgeojson, highcharter, htmlwidgets)

```


## Pre-processing

Create a geometry column and drop unneccessary information
```{r}
# Importing the data.
data <- read_csv("data/coffee_arabica_LongLat.csv")


# remove farms without coordinates. The locations are annotated comma seperated.
data <- data %>% drop_na(location_coordinates)


# Create longitude and latitude column seperately (X and Y)
coordinates <- data.frame(do.call("rbind", strsplit(as.character(data$location_coordinates), ",", fixed = TRUE)))

# Create appropriate column names and change to numeric
coordinates <- coordinates %>% 
  rename(
    Y = X1,
    X = X2
    ) %>% 
  mutate(
    X = as.numeric(as.character(X)),
    Y = as.numeric(as.character(Y))
  )

# Combine the new coordinate columns with the data.
data <- cbind(data, coordinates)
rm(coordinates)

# Make geometry point column. We choose the crs to be 4326.
data <- st_as_sf(data, coords = c("X", "Y"), remove = FALSE, 
    crs = 4326, agr = "constant")

# check crs
st_crs(data)

# Only keep relevant columns
data <- data[c("X", "Y", "Region","Country.of.Origin","Farm.Name", "Processing.Method", "unit_of_measurement", "altitude_low_meters", "altitude_high_meters", "altitude_mean_meters", "Aroma", "Flavor", "Aftertaste", "Acidity", "Body", "Balance", "Uniformity", "Sweetness")]

# Remove one observation where all scores are 0.
data <- data %>% filter(Acidity != 0)

```

A quick overview of the data
```{r}
# Summary of the data
summary(data)

# Mapview to check correctness of lat and long. There seems to be some false datapoints in the oceans and outside coffee regions.
mapview(data)

```

Clean up columns with altitude and processing methods
```{r}
# Removing altitudes that are clearly artefacts (coffee cannot grow 300 meters above sea level)
data$altitude_mean_meters <- ifelse(0 < data$altitude_mean_meters & 3000 > data$altitude_mean_meters, data$altitude_mean_meters, NA)

# Gather the processing methods in 4 overall categories:

# semi
data$Processing.Method <- ifelse(data$Processing.Method == "Pulped natural / honey" | data$Processing.Method == "Semi-washed / Semi-pulped", "semi", data$Processing.Method)
# washed
data$Processing.Method <- ifelse(data$Processing.Method == "Washed / Wet", "washed", data$Processing.Method)
# natural
data$Processing.Method <- ifelse(data$Processing.Method == "Natural / Dry", "natural", data$Processing.Method)
# other
data$Processing.Method <- ifelse(data$Processing.Method == "Other", "other", data$Processing.Method)


# change column into factor and create a more appropriate name for it
data <- data %>%  mutate(
  Processing.Method = as.factor(Processing.Method)
)
colnames(data)[which(names(data) == "Processing.Method")] <- "processing_method"

# the processing type 'washed' is held as the reference category
data$processing_method <- relevel(data$processing_method, ref = "washed")


# Overview

# histogram
hist(data$altitude_mean_meters, xlab = "Mean Altitude (m)")
# Seems fairly normally distributed (That is an underlying assumption when we are feeding parameters into linear regression models)

# Number in each processing category
data %>% group_by(processing_method) %>% summarise(
  sum = n()
) # Very unevenly distributed. "washed" is by far the most common processing method. This should be kept in mind.

```


## Moran´s I

In order to assess spatial auto-correlation on various coffee characteristics, we employ a Moran's I. 
Since Moran's I is not really robust (as it is very sensitive to single values), we run Moran's I monte Carlo tests to assess how significant any potential correlations actually are. In that way, we can investigate how likely the data is given the assumption that the distribution is random (no spatial auto-correlation). Under a significance level of 0.05, this nul-hypothesis will be rejected.

### Get neighbors and geographical weights
```{r}

# get neighbors (we define neighbors as within an Euclidean distance)
distance <- 200

# The function dnearneigh() identifies neighbours of every point by Euclidean distance between lower and upper bounds.
coords <- data$geometry
neighbors <- dnearneigh(coords, 0, distance, row.names=NULL)

# Summary information about neighbors
summary(neighbors, coords)


## Plot neighbor links

# Import world polygons
world <- st_read("data/world_shape_file/TM_WORLD_BORDERS_SIMPL-0.3.shp")

# Check that the crs is the same
st_crs(world) == st_crs(data)

# Show the connections
plot(world$geometry); plot(neighbors, coords, col = "red",add = TRUE)


# Using nb2listw() to create a listw (list of weights) object - a matrix of geographical weights.
weight_matrix <- nb2listw(neighbors, glist=NULL, style="W", zero.policy=TRUE)

```

### Run Moran's I Monte Carlo

A permutation test for Moran’s I statistic calculated by using nsim random permutations of the data for
the given spatial weighting scheme, to establish the rank of the observed statistic in relation to the
'nsim' simulated values.
```{r}
# Number of simulations
nsim <- 999
set.seed(1234)

```

__Balance__ 
```{r}
# Run Moran's I
mc_balance <- moran.mc(data$Balance, weight_matrix, nsim, zero.policy=TRUE, alternative="greater",
na.action=na.fail, spChk=NULL, return_boot=FALSE, adjust.n=TRUE)

# Results
mc_balance

# Plot
# In its default setting, the plot shows a linear fit through the point cloud. The slope of this line corresponds to Moran’s I
moran.plot(data$Balance, weight_matrix, zero.policy=FALSE, spChk=NULL,
xlab="Balance Score", ylab=NULL, quiet=NULL, plot=TRUE)
plot(mc_balance)

# plot the relationship between balance and its spatially lagged counterpart.
inc.lag <- lag.listw(weight_matrix, data$Balance)
plot(inc.lag ~ data$Balance, pch=16, asp=1., xlab = "Balance", ylab = "Spatially lagged neighbors")
M1 <- lm(inc.lag ~ data$Balance)
abline(M1, col="blue") # The slope of the line is the Moran’s I coefficient.

# We see spatial structure when it comes to balance

```

__Acidity__ 
```{r}
# Run Moran's I
mc_acidity <- moran.mc(data$Acidity, weight_matrix, nsim, zero.policy=TRUE, alternative="greater",
na.action=na.fail, spChk=NULL, return_boot=FALSE, adjust.n=TRUE)

# Results
mc_acidity

# Plot
# In its default setting, the plot shows a linear fit through the point cloud. The slope of this line corresponds to Moran’s I
moran.plot(data$Acidity, weight_matrix, zero.policy=FALSE, spChk=NULL, labels=NULL,
xlab="Acidity Score", ylab=NULL, quiet=NULL, plot=TRUE)
plot(mc_balance)

# We see spatial structure when it comes to acidity - the estimate is however not as big large with balance

# plot the relationship between balance and its spatially lagged counterpart.
inc.lag <- lag.listw(weight_matrix, data$Acidity)
plot(inc.lag ~ data$Acidity, pch=16, asp=1., xlab = "Aciditye", ylab = "Spatially lagged neighbors")
M1 <- lm(inc.lag ~ data$Acidity)
abline(M1, col="blue") # The slope of the line is the Moran’s I coefficient.


```

__Aroma__ 
```{r}
# Run Moran's I
mc_aroma <- moran.mc(data$Aroma, weight_matrix, nsim, zero.policy=TRUE, alternative="greater",
na.action=na.fail, spChk=NULL, return_boot=FALSE, adjust.n=TRUE)

# Results
mc_aroma

# Plot
# In its default setting, the plot shows a linear fit through the point cloud. The slope of this line corresponds to Moran’s I
moran.plot(data$Aroma, weight_matrix, zero.policy=FALSE, spChk=NULL, labels=NULL,
xlab="Aroma Score", ylab=NULL, quiet=NULL, plot=TRUE)
plot(mc_aroma)

# We see spatial structure when it comes to aroma as well.

# plot the relationship between balance and its spatially lagged counterpart.
inc.lag <- lag.listw(weight_matrix, data$Aroma)
plot(inc.lag ~ data$Aroma, pch=16, asp=1., xlab = "Aroma", ylab = "Spatially lagged neighbors")
M1 <- lm(inc.lag ~ data$Aroma)
abline(M1, col="blue") # The slope of the line is the Moran’s I coefficient.

```

## Altitude 

As altitude is know for being an essential spatial component when it comes to coffee production, we will as well predict the taste characteristics from altitude. Additionally, we add the way in which the beans hve been processed as a predictor.

```{r}
# Balance
model_balance <- lm(Balance ~ altitude_mean_meters + processing_method, data = data)
summary(model_balance)
plot(model_balance)

# Acidity
model_acidity <- lm(Acidity ~ altitude_mean_meters + processing_method, data = data)
summary(model_acidity)

# Aroma
model_aroma <- lm(Aroma ~ altitude_mean_meters + processing_method, data = data)
summary(model_aroma)


```


## Visualizations

### Mapping locations

Mapping the location of coffee growing farms around the world. One can click on each farm and read its specific scores on various parameters.
```{r}
#creating a map with leaflet that displays the coffee farms + layers to manipulate
basemap <- leaflet() %>%
  # add different provider tiles
  addProviderTiles("OpenStreetMap",group = "OpenStreetMap") %>%
  addProviderTiles("Stamen.Toner",group = "Stamen.Toner" ) %>%
  addProviderTiles("Stamen.Terrain", group = "Stamen.Terrain") %>%
  addProviderTiles("Esri.WorldStreetMap", group = "Esri.WorldStreetMap") %>%
  addProviderTiles("Wikimedia",group = "Wikimedia" ) %>%
  addProviderTiles("CartoDB.Positron",group = "CartoDB.Positron") %>%
  addProviderTiles("Esri.WorldImagery",group = "Esri.WorldImagery") %>%
# add a layers control
  addLayersControl(baseGroups = c(
      "OpenStreetMap", "Stamen.Toner",
      "Stamen.Terrain", "Esri.WorldStreetMap",
      "Wikimedia", "CartoDB.Positron", "Esri.WorldImagery"),
    # position it on the topleft
    position = "topleft")

#adding the pins 
icon.fa <- makeAwesomeIcon(
  icon = "coffee", markerColor = "green",
  library = "fa",
  iconColor = "black"
)

#merging the map and icons
map_1 <- basemap %>%
  addAwesomeMarkers(
    lat = data$Y,
    lng =data$X,
    label = data$Region,
    icon = icon.fa
  )

#display
map_1

saveWidget(map_1, file="maps/farms_map.html")


```

### Mapping tendencies by country

Building an interactive map - one for aroma, one fore acidity and one for balance. They illustrate an average by country to enable a quick overview of general trends by country.
```{r}

#necessary preprocessing 
#renaming country names so that they match the shapes of the countries - for plotting 
data$Country.of.Origin[data$Country.of.Origin == "United States (Hawaii)"] <- "United States of America"
data$Country.of.Origin[data$Country.of.Origin == "United States (Puerto Rico)"] <- "United States of America"
data$Country.of.Origin[data$Country.of.Origin == "United States"] <- "Colombia" # there has been a mistake and all beans marked as coming from the USA were in fact from Columbia 
data$Country.of.Origin[data$Country.of.Origin == "Tanzania, United Republic Of"] <- "Tanzania"


#aggregating scores by countries + rounding to 2 decimal places 
# mean aroma by country+  renaming columns
mean_aroma_by_country <- aggregate(data$Aroma,by=list(data$Country.of.Origin), mean) 
colnames(mean_aroma_by_country)[1] <- "country"
mean_aroma_by_country$x <- round(mean_aroma_by_country$x,2)

# mean acidity by country+  renaming columns +rounding to 2 decimal places 
mean_acidity_by_country <- (aggregate(data$Acidity,by=list(data$Country.of.Origin), mean))
colnames(mean_acidity_by_country)[1] <- "country"
mean_acidity_by_country$x <- round(mean_acidity_by_country$x,2)


# mean balance by country+  renaming columns +rounding to 2 decimal places 
mean_balance_by_country <- aggregate(data$Balance,by=list(data$Country.of.Origin), mean) 
colnames(mean_balance_by_country)[1] <- "country"
mean_balance_by_country$x <- round(mean_balance_by_country$x,2)

# Find min and max scores by country to create an appropriate scale
data %>% group_by(country)


#building interactive maps 
#Aroma
aroma_map <- highchart() %>%
  hc_add_series_map(
    worldgeojson,mean_aroma_by_country, value = "x", joinBy =c("name",'country'),
    name = "Mean score"
    )  %>% 
  hc_colorAxis( min = 7, max = 8.4,minColor="#F1B6DA",maxColor="#CA0020") %>% 
  hc_title(text = "World Map") %>% 
  hc_subtitle(text = "Coffee Aroma")

# Run map
aroma_map
# Save map
saveWidget(aroma_map, "maps/aroma_map.html", selfcontained = TRUE)


#Acidity
acidity_map <- highchart() %>%
  hc_add_series_map(
    worldgeojson,mean_acidity_by_country, value = "x", joinBy =c("name",'country'),
    name = "Mean score"
    )  %>% 
  hc_colorAxis( min = 6.6, max = 8.4,minColor="#EFFF7F",maxColor="#E66101") %>% 
  hc_title(text = "World Map") %>% 
  hc_subtitle(text = "Coffee Acidity")

# run map
acidity_map
# Save map
saveWidget(acidity_map, "maps/acidity_map.html", selfcontained = TRUE)


#Balance
balance_map <- highchart() %>%
  hc_add_series_map(
    worldgeojson,mean_balance_by_country, value = "x", joinBy =c("name",'country'),
    name = "Mean score"
    )  %>% 
  hc_colorAxis( min = 7, max = 8.3,minColor="#B8E186",maxColor="#018571") %>% 
  hc_title(text = "World Map") %>% 
  hc_subtitle(text = "Coffee Balance")

# run map
balance_map
# Save map
saveWidget(balance_map, "maps/balance_map.html", selfcontained = TRUE)

```


















